Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-3\baselineskip}

### 2017-01-08 paper/ms.tex
words about asterosesmology

{\item For $Q = 1/\sqrt{2}$, \eq{sho-psd} simplifies to
\begin{eqnarray}
S(\omega) = \sqrt{\frac{2}{\pi}}\,\frac{S_0}{(\omega/\omega_0)^4+1} \quad.
\end{eqnarray}
A model like this has been used to model the granulation is Solar data
\citep{Michel:2009}}, which has a kernel of
\begin{equation}
k(\tau) = S_0\,\omega_0\,e^{-\frac{1}{\sqrt{2}}\,\omega_0\,\tau}\,
    \cos{\left(\frac{\omega_0\,\tau}{\sqrt{2}}-\frac{\pi}{4}\right)}.
\end{equation}

@@

\section{Examples}

\subsection{Simulated data}

@@

\subsection{Real data}

\begin{enumerate}
\item AGN
\item RR Lyrae
\item Kepler with simulated transit
\item Asteroseismic target
\item Stellar rotational variability
\end{enumerate}

>>>>

{\item For $Q = 1/\sqrt{2}$, \eq{sho-psd} simplifies to
\begin{eqnarray}\eqlabel{granulation-psd}
S(\omega) = \sqrt{\frac{2}{\pi}}\,\frac{S_0}{(\omega/\omega_0)^4+1} \quad.
\end{eqnarray}
A model like this has been used to model the granulation is Solar data
\citep{Michel:2009}}, which has a kernel of
\begin{equation}
k(\tau) = S_0\,\omega_0\,e^{-\frac{1}{\sqrt{2}}\,\omega_0\,\tau}\,
    \cos{\left(\frac{\omega_0\,\tau}{\sqrt{2}}-\frac{\pi}{4}\right)}.
\end{equation}

@@

\section{Examples with simulated data}

@@

\section{Examples with real data}

\subsection{Asteroseismic oscillations}

The asteroseismic oscillations of thousands of stars were measured using light
curves from the \kepler\ mission (\dfmtodo{cite}) and asteroseismology is a
key science driver for many of the upcoming large scale photometric surveys
(\dfmtodo{cite}).
Most asteroseismic analyses have been limited to relatively high
signal-to-noise oscillations because the standard methods based on statistics
of the empirical periodogram of the data cannot be used to formally propagate
the measurement uncertainties to the constraints on physical parameters
(\dfmtodo{CITE}) and more sophisticated methods are computationally expensive
and they scale poorly to the current state-of-the-art datasets (\dfmtodo{CITE
Brewer etc.}).

\genrp\ alleviates these problems by providing a physically motivated
probabilistic model that can be evaluated efficiently even for large datasets.
In practice, we will model the star as a mixture of stochastically driven
simple harmonic oscillators where the amplitudes and frequencies of the
oscillations are computed using a physical model and evaluate the probability
of the observed dataset using a Gaussian Process with a PSD given by a sum of
terms given by \eq{sho-psd}.
This gives us a method of computing the likelihood function for the parameters
of the physical model in $\mathcal{O}(N)$ operations and this can be combined
with standard non-linear optimization or posterior sampling methods to infer
constraints on the parameters for the given dataset.

To demonstrate this method, we will use a very simple heuristic model based on
\dfmtodo{CITE} where the model PSD is given by a mixture of 6 components with
amplitudes and frequencies specified by $\nu_\mathrm{max}$, $\Delta \nu$, and
several nuisance parameters.
The first term is used to capture the granulation ``background'' using
\eq{granulation-psd} with two free parameters $S_g$ and $\omega_g$.
The remaining 5 terms are given by \eq{sho-psd} where $Q$ is a nuisance
parameter shared between terms and the frequencies are given by
\begin{eqnarray}
\omega_{0,\,j} = 2\,\pi\,(\nu_\mathrm{max} + j\,\Delta\nu + \epsilon)
\end{eqnarray}
and the amplitudes are given by
\begin{eqnarray}
S_{0,\,j} =
    \frac{A}{Q^2}\,\exp\left(-\frac{[j\,\Delta\nu + \epsilon]^2}{2\,W^2}\right)
\end{eqnarray}
where $j$ is an integer running from $-2$ to 2 and $\epsilon$, $A$, and $W$
are shared nuisance parameters.
This model could be easily extended to include small frequency splitting and
$\nu_\mathrm{max}$ and $\Delta \nu$ could be replaced by physical parameters
like $\log g$.

\dfmtodo{Add plots etc.}

\begin{enumerate}
\item Asteroseismic target
\item Stellar rotation variability
\item Kepler with simulated transit
\item AGN
\end{enumerate}
